---
title: "Predicting for Coursera - Practical Machine Learning"
output: html_document
---

Dividing Data into Training and Testing Data. Due to computation time, only a fraction of all data will be used.

```{r warning=FALSE}
dataLoad <- read.csv("pml-training.csv")
library("caret")

set.seed(721)
data <- dataLoad[createDataPartition(dataLoad$classe, p=0.04, list=FALSE),]
set.seed(1500)
testingData <- dataLoad[createDataPartition(dataLoad$classe, p=0.04, list=FALSE),]

```

## 1. Description of the outcome variable
```{r}
summary(dataLoad$classe)
```


Searching for important predictor variables by investigating the correlation matrix
```{r warning=FALSE}
cor.Matrix <- rep(NA, times=159)
cor.Matrix.sig <- rep(NA, times=159)
classeA <- ifelse(dataLoad$class=="A", 1, 0)
for (i in 8:159) {
cor.Matrix[i] <- ifelse(class(dataLoad[,i])=="factor", NA, cor(dataLoad[,i], classeA))
cor.Matrix.sig[i] <- ifelse(class(dataLoad[,i])=="factor", NA, ifelse(cor(dataLoad[,i], classeA)>0.02, round(cor(dataLoad[,i], classeA), digits=2),  "small"))
}

```

Important variables are contained in the following columns: 42, 44, 45, 47, 49, 64, 67, 68, 102, 154, 157, 158

## Random Forest
As there are more than two outcomes, I will try using Random forests.

```{r warning=FALSE}
library(randomForest)



trainRF <- train(data$classe ~ ., data=data[, c(42, 44, 45, 47, 49, 64, 67, 68, 102, 154, 157, 158, 160)], method="rf", prox=TRUE)


predRF <- predict(trainRF, data)
table(predRF, data$classe)
```



Misclassification error:

`r 1-sum(diag(matrix(table(predRF, data$classe), nrow=5, ncol=5)))/sum(matrix(table(predRF, data$classe), nrow=5, ncol=5)) `

## Linear Discrimination
As an alternative I will try linear discriminant analysis

```{r}

trainLDA <- train(data$classe ~ ., data=data[, c(42, 44, 45, 47, 49, 64, 67, 68, 102, 154, 157, 158, 160)], method="lda", prox=TRUE)

predLDA <- predict(trainLDA, data)
table(predLDA, data$classe)
```



Misclassification error:
`r 1-sum(diag(matrix(table(predLDA, data$classe), nrow=5, ncol=5)))/sum(matrix(table(predLDA, data$classe), nrow=5, ncol=5)) `

As the Error (in the subsample) is higher for the Linear Discriminant analysis. I´d use the Random Forest algorithm. Of course the true error would be higher for real data, but the classification is pretty successful. As a better estimation of the true estimation error, it can be calculated based on the testingData:

```{r}
predRF <- predict(trainRF, testingData)
table(predRF, testingData$classe)
1-sum(diag(matrix(table(predRF, data$classe), nrow=5, ncol=5)))/sum(matrix(table(predRF, data$classe), nrow=5, ncol=5))


predLDA <- predict(trainLDA, testingData)
table(predLDA, testingData$classe)
1-sum(diag(matrix(table(predLDA, data$classe), nrow=5, ncol=5)))/sum(matrix(table(predLDA, data$classe), nrow=5, ncol=5))
```

## Prediction for 20 cases
Finally: Predicting the 20 cases from the testing-File:

```{r}
TestData <- read.csv("pml-testing.csv")

finalprediction <- predict(trainRF, TestData)
print(finalprediction)
```

